---
title: "ESM 206 Lab 5"
author: "Allison Horst"
date: "10/28/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

library(tidyverse)
library(janitor)


```




```{r}

lobster_traps_tidy <- read_csv(here::here("data","lobster_traps.csv")) %>%
  janitor::clean_names() 
lobster_tidy <- read_csv(here::here("data","lobster_abundance.csv")) %>%
  janitor::clean_names() %>%
  mutate(size_mm = replace(size_mm, size_mm == -99999, NA))

view(lobster_traps_tidy)
view(lobster_tidy)

```



```{r}

# ggplot(data = lobster_tidy, aes(x = site, y = size_mm)) + 
#   geom_violin(aes(color = site), 
#               alpha = 0.3)
# 
# 
# ggplot(data = lobster_tidy, aes(x = size_mm)) + 
#   geom_histogram(aes(fill = site)) + 
#   
#
ggplot(data = lobster_tidy, aes(sample = size_mm)) + 
  geom_qq(size = 0.5) +
  facet_wrap(~site, scales = "free")
  

```

Use `lubridate` to parse dates and times: 
NOTE - a lot of lubridate functions are highly overloaded against other core libraries... good idea to explicitly specify the namespace here!

```{r}

lobster_date <- lobster_tidy %>%
  mutate(
    date_new = lubridate::mdy(date)   # MDY for Month / Day / Year format
  )

```

Parse the date column to get the different pieces (month, year) separated

```{r}

lobster_parse_date <- lobster_date %>%
  mutate(
    obs_month = lubridate::month(date_new, label = TRUE),   # the TURE flag makes the returns be string abbreviations, not month numbers 
    obs_year = lubridate::year(date_new)
  )

lobster_parse_date

```

Count lobsters by different groupings....

Let's say I want to count the number of lobsters by year and month

```{r}

lobster_ym <- lobster_parse_date %>%
  count(obs_year, obs_month)

lobster_ym

lobster_ysite <- lobster_parse_date %>%
  count(obs_year, site)

lobster_ysite

lobster_site <- lobster_parse_date %>%
  count(site)

lobster_site

lobster_summary <- lobster_parse_date %>%
  group_by(site) %>%
  summarize(
    mean_size = mean(size_mm, na.rm = TRUE),
    sd_size = sd(size_mm, na.rm = TRUE),
    sample_size = n()
  )

lobster_summary

```

Confidence intervals and t-tests

Use the `t.test()` function to find confidence intervals and perform t-tests

```{r}

ivee_lobsters <- lobster_tidy %>%
  filter(site == "IVEE") %>%
  pull(size_mm)  # returns just a vector for the selected variable  

ivee_lobsters

t.test(ivee_lobsters)

min(ivee_lobsters, na.rm = TRUE)

```

Two-sample t-test:
Is there a significant difference in mean lobster lengths between Naples Reef and Mohawk Reef?

```{r}

napl_lobsters <- lobster_tidy %>%
  filter(site == "NAPL") %>%
  pull(size_mm)  # returns just a vector for the selected variable  
mohk_lobsters <- lobster_tidy %>%
  filter(site == "MOHK") %>%
  pull(size_mm)  # returns just a vector for the selected variable  

test_val <- t.test(napl_lobsters, mohk_lobsters, paired = FALSE)

```

Mean lobster size differed significantly between Mohawk and Naples Reefs (t(`r round(test_val$parameter,2)`) = `r round(test_val$statistic,2)`, p = `r test_val$p.value`)


Maybe an easier way...

```{r}
# Filter to retain only two groups
lobster_2sample <- lobster_tidy %>%
  filter(site %in% c("NAPL", "MOHK"))

ttest_2 <- t.test(size_mm ~ site, data = lobster_2sample)

ttest_2

```

Make a geom_tile heatmap

```{r}

ggplot(data = lobster_ysite, aes(x = obs_year, y = site)) + 
  geom_tile(aes(fill = n))

```








